---
title: "STA 5900 Project 3 Logistic Prediction"
author: "Theodore De Santos"
date: "April 2022"
output:
    pdf_document:
      df_print: kable
---

```{r global_options, include=FALSE}
# these are some optional settings that will change how some features look
# you do not need to change them.
knitr::opts_chunk$set(out.width = "50%", out.height="50%", fig.align="center", warning=FALSE, message=FALSE)
```

```{r,include=FALSE}
library(caret)
library(glmnet)
library(leaps)
library(DescTools)
library(stringr)
library(scales)
library(dplyr)
library(rpart)
library(rpart.plot)
library(randomForest)
library(kernlab)
library(pracma)
library(questionr)
```

####################################### PLEASE CHECK IF THE DATASETS ARE THE SAME
The code below does some preparations with the dataset.  

```{r}
titanic <- read.csv('train.csv')

titanic.test <- read.csv('test.csv')

titanic.test$Survived <- rep('NA',nrow(titanic.test))

titanicPlus <- rbind(titanic, titanic.test)


# ---------- Preparing Variables For Models ----------

# Make Variables categorical

titanicPlus$Sex <- as.factor(titanicPlus$Sex)
titanicPlus$Survived <- as.factor(titanicPlus$Survived)
levels(titanicPlus$Survived) <- c('No', 'Yes', 'NA')
titanicPlus$Embarked <- as.factor(titanicPlus$Embarked)
titanicPlus$Pclass <- as.factor(titanicPlus$Pclass)

# Fare price for Passenger 1044 is blank
titanicPlus$Fare[titanicPlus$PassengerId==1044] <- 7.5

# Split Name into First and Last

split.name <- strsplit(titanicPlus$Name, ',')

Lastname <- sapply(split.name, function(x) x[1])
Firstname <- sapply(split.name, function(x) x[2])

titanicPlus$Lastname <- Lastname

# Adding Number of Occurance of Last Name to each person

Lastname.Occur <- Freq(titanicPlus$Lastname)

titanicPlus$Occur <- rep(1,length(Lastname))
for (i in Lastname.Occur$level) {
  titanicPlus$Occur[which(Lastname==i)] <- Lastname.Occur[Lastname.Occur$level==i,]$freq
}

# Split Prefix of First Name

split.prefix <- strsplit(Firstname, '. ')

titanicPlus$Prefix <- sapply(split.prefix, function(x) x[1])

titanicPlus$Prefix <- as.factor(titanicPlus$Prefix)

# Separate Cabin Letter

split.cabin <- strsplit(titanicPlus$Cabin, '')

titanicPlus$Letter <- sapply(split.cabin, function(x) x[1])

# Adding Cabin Letter to Family members

Letters.notNA <- titanicPlus[-which(is.na(titanicPlus$Letter)),]

for (i in Letters.notNA$Lastname[Letters.notNA$Occur>1]) {
  titanicPlus$Letter[which(Lastname==i)] <- unique(Letters.notNA$Letter[Letters.notNA$Lastname==i])[1]
}

titanicPlus$Letter <- as.factor(titanicPlus$Letter)



# ------ Creating CSV for the new dataset ------

write.csv(titanicPlus,'new_train.csv',row.names=FALSE)
```

The following code will load in the new data set and set the factor variables 

```{r}
titanicPlus <- read.csv('new_train.csv')

titanicPlus$Survived[892:nrow(titanicPlus)] <- 'NA'

# ------ Changing variable to categorical ------

titanicPlus$Sex <- as.factor(titanicPlus$Sex)
titanicPlus$Survived <- as.factor(titanicPlus$Survived)
titanicPlus$Embarked <- as.factor(titanicPlus$Embarked)
titanicPlus$Pclass <- as.factor(titanicPlus$Pclass)
titanicPlus$Letter <- as.factor(titanicPlus$Letter)
titanicPlus$Prefix <- as.factor(titanicPlus$Prefix)
```

An issue has been that there are missing values in the variables that we need access to, so Connor has used the following code to fill in these missing values

```{r}
# ---------- Model for Filling in Cabin Letters ----------

# Random Forest Model for predicting Cabin Letters
loop.vector = c(1:5)
for (i in loop.vector) {
  Cabin.Forest <- randomForest(Letter~Fare+
                               Survived+
                               SibSp+
                               Parch+
                               SibSp*Parch+
                               Survived*SibSp+
                               Pclass*Survived+
                               Pclass+
                               Occur,
                             method='class',
                             na.action=na.omit,
                             data=titanicPlus,
                             proximity=TRUE
                             )

if (length(which(is.na(titanicPlus$Letter)))>0){
  cabin.na <-which(is.na(titanicPlus$Letter))
}
}

predict.Letter <- predict(Cabin.Forest,titanicPlus[cabin.na,],
                         type='class')

titanicPlus$Letter[cabin.na] <- predict.Letter


# Lm Model for Predicting Age

  # Creating Age.Na column for Ages not entered

if (length(which(is.na(titanicPlus$Age)))>0){
  Age.na <- which(is.na(titanicPlus$Age))
}

titanicPlus$Age.Na <- 0
titanicPlus$Age.Na[which(is.na(titanicPlus$Age))] <- 1

Age.lm <- lm(Age~Sex+
               SibSp+
               Parch+
               Prefix*Fare+
               SibSp*Parch+
               Pclass*Fare+
               Pclass*Parch+
               Pclass+
               Occur+
               Occur*SibSp+
               Survived+
               Embarked+
               Letter+
               Prefix+
               Fare,
             na.action=na.omit,
             data=titanicPlus
             )

predict.Age <- predict(Age.lm,titanicPlus[-Age.na,])

RMSE.age <- sqrt(mean((predict.Age - 
                         titanicPlus$Age[-Age.na])^2)); RMSE.age

predict.Age <- predict(Age.lm,titanicPlus[Age.na,])

titanicPlus$Age[Age.na] <- predict.Age
```

This code will establish the new training set that will be used.  


### Analysis

This block will first make the proper training and testing set

NOTE:  THE AIC AND Accuaracy (ACC) values may be inaccurate if the dataset manipulation above needed fixing


```{r}
trainSet <- titanicPlus[-which(titanicPlus$Survived=='NA'),]
if (length(which(titanicPlus$Survived=='NA'))<1) {
  trainSet = titanicPlus
}

trainSet$Survived <- as.numeric(trainSet$Survived)
trainSet$Survived <- as.factor(trainSet$Survived)

levels(trainSet$Survived) <- c('No', 'Yes')
```
###############################


This analysis will focus on fitting logistic regression models.

What will first be done is fitting a logisitic regression model with all variables

```{r}
glm.fitall <- glm(Survived ~ .-Name-Ticket-Lastname-PassengerId-Cabin, data = trainSet,family=binomial)
phat <- predict(glm.fitall, data=trainSet,type = "response")
yhat <- rep("No", length.out=nrow(trainSet))
yhat[which(phat>0.5)] <- "Yes"
confusionMatrix(trainSet$Survived, yhat, positive='Yes')
```
AIC:729.17
ACC: 0.8361


```{r}
glm.fitall <- glm(Survived ~ .-Name-Ticket-Lastname-PassengerId-Cabin-Parch, data = trainSet,family=binomial)
phat <- predict(glm.fitall, data=trainSet,type = "response")
yhat <- rep("No", length.out=nrow(trainSet))
yhat[which(phat>0.5)] <- "Yes"
confusionMatrix(trainSet$Survived, yhat, positive='Yes')
```

AIC 727.86
ACC: 0.8383


```{r}
glm.fitall <- glm(Survived ~ .-Name-Ticket-Lastname-PassengerId-Cabin-Parch-Age.Na, data = trainSet,family=binomial)
phat <- predict(glm.fitall, data=trainSet,type = "response")
yhat <- rep("No", length.out=nrow(trainSet))
yhat[which(phat>0.5)] <- "Yes"
confusionMatrix(trainSet$Survived, yhat, positive='Yes')
```

AIC 725.86
ACC: 0.8372

```{r}
glm.fitall <- glm(Survived ~ .-Name-Ticket-Lastname-PassengerId-Cabin-Parch-Age.Na-Sex, data = trainSet,family=binomial)
phat <- predict(glm.fitall, data=trainSet,type = "response")
yhat <- rep("No", length.out=nrow(trainSet))
yhat[which(phat>0.5)] <- "Yes"
confusionMatrix(trainSet$Survived, yhat, positive='Yes')
```

AIC 725.33
ACC: 0.8383

```{r}
glm.fitall <- glm(Survived ~ .-Name-Ticket-Lastname-PassengerId-Cabin-Parch-Age.Na-Sex-Embarked, data = trainSet,family=binomial)
phat <- predict(glm.fitall, data=trainSet,type = "response")
yhat <- rep("No", length.out=nrow(trainSet))
yhat[which(phat>0.5)] <- "Yes"
confusionMatrix(trainSet$Survived, yhat, positive='Yes')
```

AIC 724.76
ACC: 0.8429

```{r}
glm.fitall <- glm(Survived ~ .-Name-Ticket-Lastname-PassengerId-Cabin-Parch-Age.Na-Sex-Embarked+Age*Letter, data = trainSet,family=binomial)
phat <- predict(glm.fitall, data=trainSet,type = "response")
yhat <- rep("No", length.out=nrow(trainSet))
yhat[which(phat>0.5)] <- "Yes"
table(yhat,trainSet$Survived)
summary(glm.fitall)
```

AIC 725.58
ACC: 0.8462